The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-12 manuscript/manuscript.Rnw
move figures to be closer discussion point

@@

@@

\begin{figure}
<<tmsb1.receptor.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="TMSB10" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{\textit{TMSB1}}}
\label{fig:tmsb1_evolution_tree}
\end{figure}

The elephant is strikingly the only mammal analyzed that shows high
expression of the primarily pituitary hormone prolactin (PRL) in the
placenta. All afrotherians yet analyzed contain only one prolactin
gene, and prolactin has undergone adaptive evolution on the elephant
lineage \cite{Wallis2009:ProlactininAfrotheriacharacterization}.
\citet{Yamamoto.ea2011:PlacentationinAfricanelephant} noted the
presence of a prolactin in the trophoblast of the elephant placenta
using immunohistological staining and human derived antibodies. A
previous analysis of the elephant placenta also detected the
expression of prolactin
\cite{Hou.ea2012:Elephanttranscriptomeprovidesinsights}. Here we
confirm that the sole prolactin in the elephant is not only expressed
but highly expressed in the placenta, and likely functions both in the
maintenance of the corpus luteum and as a placental lactogen
\cite{Yamamoto.ea2011:PlacentationinAfricanelephant}. \textit{PRL} is
not expressed in the opossum, even though there is evidence that it is
expressed in the placenta of another marsupial, the tammar wallaby
\cite{Menzies.ea2011:Placentalexpressionofpituitary}. This suggests
that placental prolactin expression in marsupials may be as diverse as
in eutherian mammals. Although, \textit{PRL} is not expressed to any
great degree in the placenta other mammals, we see very high
expression of genes which resulted from duplications of \textit{PRL}
in the \mouse, \galili, \carmeli, and \cow
(\cref{fig:prolactin_evolution_tree}).

>>>>

\subsubsection{Estrogen and Progesterone}

@@

\subsubsection{Prolactin}

The elephant is strikingly the only mammal analyzed that shows high
expression of the primarily pituitary hormone prolactin (\textit{PRL}) in the
placenta. All afrotherians yet analyzed contain only one prolactin
gene, and prolactin has undergone adaptive evolution on the elephant
lineage \cite{Wallis2009:ProlactininAfrotheriacharacterization}.
\citet{Yamamoto.ea2011:PlacentationinAfricanelephant} noted the
presence of a prolactin in the trophoblast of the elephant placenta
using immunohistological staining and human derived antibodies. A
previous analysis of the elephant placenta also detected the
expression of prolactin
\cite{Hou.ea2012:Elephanttranscriptomeprovidesinsights}. Here we
confirm that the sole prolactin in the elephant is not only expressed
but highly expressed in the placenta, and likely functions both in the
maintenance of the corpus luteum and as a placental lactogen
\cite{Yamamoto.ea2011:PlacentationinAfricanelephant}. \textit{PRL} is
not expressed in the opossum, even though there is evidence that it is
expressed in the placenta of another marsupial, the tammar wallaby
\cite{Menzies.ea2011:Placentalexpressionofpituitary}. This suggests
that placental prolactin expression in marsupials may be as diverse as
in eutherian mammals. Although, \textit{PRL} is not expressed to any
great degree in the placenta other mammals, we see very high
expression of genes which resulted from duplications of \textit{PRL}
in the \mouse, \galili, \carmeli, and \cow
(\cref{fig:prolactin_evolution_tree}).

@@

\subsubsection{\ac{GH} family}
